Missing the Point\(\colon\) Non-Convergence in Iterative Imputation Procedures

### 2020-04-24 3.Thesis/1.SimulationStudy/Writeup/manuscript.Rmd
Placed new figures in manuscript

```{r non-conv, out.width='.49\\linewidth', fig.width=3.5, fig.height=3,fig.show='hold',fig.cap="Pathological non-convergence versus typical convergence."}
load("./images/nonc.Rdata")
load("./images/conv.Rdata")
par(mar = c(2, 2, 0.1, 0.1))
plot(nonc)
plot(conv)

@@

>>>>

```{r non-conv, out.width='.49\\linewidth', fig.width=3.5, fig.height=3,fig.show='hold',fig.cap="Typical convergence versus pathological non-convergence."}
load("./images/nonc_plot.Rdata")
load("./images/conv_plot.Rdata")
par(mar = c(2, 2, 0.1, 0.1))
plot(conv)
plot(nonc)

@@

```{r diagnostics, fig.cap="Convergence diagnostics pathological non-convergence versus typical convergence."}
library(ggplot2)
library(patchwork)
theme_update(
  plot.title = element_text(hjust = 0.5),
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black"),
  legend.key = element_blank(),
  legend.position = "bottom",
  legend.margin = margin(0,0,0,0),
  # adjust text size
  text = element_text(size = 8),
  title = element_text(size = 8)
  # axis.title = element_text(size = 8),
  # axis.text = element_text(size = 7),
  # legend.text = element_text(size = 7)
)

load("./images/diagnostics_plot.Rdata")
diagnostics_plot
```

### 2020-04-24 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex
Placed new figures in manuscript

\textbf{Dear Daniela, I have just implemented some major changes in my
Methods and Results sections. Therefore, these are kind of a mess.
Please do not waste any time on spelling/wording there, I will continue
improving it. On the other hand, if you have general advice, please tell
me because I'm happy to move thngs around still! Figures in the Results
are currently missing, but will help in understanding the paper. Sorry
that you won't have that now! Also, things written in bold are reminders
for myself or things that I am unsure off still. Thanks already! -
Hanne}

@@

Missing data problems are ubiquitous and pervasive in the social and
behavioral sciences. If a dataset contains just one missing value for a
variable of interest, statistical inferences are undefined and will not
produce any result. Incomplete observations are therefore ignored by
default in many statistical packages (i.e., list-wise deletion is
employed). Unfortunately, this \emph{ad hoc} solution may yield wildly
invalid results \citep{buur18}. An alternative is to `impute' (i.e.,
fill in) every missing value in an incomplete dataset. With imputation
techniques, one or several completed datasets are created, on which
statistical inferences can be performed. The case with several completed
datasets is known as `multiple imputation' \citep[MI;][]{rubin76}, and
requires an additional step to pool the results across imputations
\textbf{see figure 1??}.

\textbf{Move this after notation??} Figure \ref{fig:diagram} provides an
overview of the steps involved with MI---from incomplete data, to \(M\)
completed datasets, to \(M\) estimated quantities of interest
(\(\hat{Q}\)s), to a single pooled estimate \(\bar{Q}\). Missing data in
\(y\) is `imputed' (i.e., filled in) \(M\) times. The imputed data
(\(y_{imp}\)) is combined with the observed data (\(y_{obs}\)) to create
\(M\) completed datasets. On each completed dataset, the analysis of
scientific interest (the `substantive model', or `complete data model')
is performed. The quantity of scientific interest (e.g., a regression
coefficient) is denoted with \(Q\). Since \(Q\) is estimated on each
completed dataset, \(M\) separate \(\hat{Q}\)-values are obtained. These
\(M\) values are combined into a single pooled estimate \(\bar{Q}\).
\textbf{ALSO: make clear that iteration happens within the first step.}

@@

\caption{Pathological non-convergence versus typical convergence.}\label{fig:non-conv}
\end{figure}

>>>>

@@

Missing data problems are ubiquitous and pervasive in the social and
behavioral sciences. If a dataset contains just one missing value for a
variable of interest, statistical inferences are undefined and will not
produce any result. To circumvent this, incomplete observations are
ignored by default in many statistical packages (i.e., list-wise
deletion is employed). Unfortunately, this \emph{ad hoc} solution may
yield wildly invalid results \citep{buur18}. An alternative is to
`impute' (i.e., fill in) every missing value in an incomplete dataset.
With imputation techniques, one or several completed datasets are
created, on which statistical inferences can be performed. The case with
several completed datasets is known as `multiple imputation'
\citep[MI;][]{rubin76}, and requires an additional step to pool the
results across imputations \textbf{see figure 1??}.

\textbf{Move this after notation??} Figure \ref{fig:diagram} provides an
overview of the steps involved with MI---from incomplete data, to \(M\)
completed datasets, to \(M\) estimated quantities of interest
(\(\hat{Q}\)s), to a single pooled estimate \(\bar{Q}\). Missing data in
\(y\) is `imputed' (i.e., filled in) \(M\) times. The imputed data
(\(y_{imp}\)) is combined with the observed data (\(y_{obs}\)) to create
\(M\) completed datasets. On each completed dataset, the analysis of
scientific interest (the `substantive model', or `complete data model')
is performed. The quantity of scientific interest (e.g., a regression
coefficient) is denoted with \(Q\). Since \(Q\) is estimated on each
completed dataset, \(M\) separate \(\hat{Q}\)-values are obtained. These
\(M\) values are combined into a single pooled estimate \(\bar{Q}\).
\textbf{ALSO: make clear that iteration happens within the first step,
between incomplete and completed data.}

@@

\caption{Typical convergence versus pathological non-convergence.}\label{fig:non-conv}
\end{figure}

{\centering \includegraphics{manuscript_files/figure-latex/diagnostics-1}

\caption{Convergence diagnostics pathological non-convergence versus typical convergence.}\label{fig:diagnostics}
\end{figure}
